
```{r}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
setwd("C:/RNA_seq_DEA")
Coinfection.targets<-read.delim("./data/fileDesc.txt")
rownames(Coinfection.targets)<-c("Ha1","Ha2","Ha3","Ctr1","Ctr2","Ctr3")
library(edgeR)
Coinfection.orig <- readDGE(Coinfection.targets, header=F)
dim(Coinfection.orig)
head(Coinfection.orig)
Coinfection.rawCount <- Coinfection.orig$count
dim(Coinfection.rawCount)
head(Coinfection.rawCount)
sampletype <- factor(c(rep("Ha",3), rep("Ctr", 3)))
meta <- data.frame(sampletype, row.names = colnames(Coinfection.orig$count))
colnames(Coinfection.orig$count)
rownames(meta)
all(colnames(Coinfection.orig$count) %in% rownames(meta))
library(DESeq2)
dds <- DESeqDataSetFromMatrix(Coinfection.orig, colData = meta, design = ~ sampletype)
head(counts(dds))
dds <- estimateSizeFactors(dds)
sizeFactors(dds)
normalized_counts <- counts(dds, normalized=TRUE)
write.csv(normalized_counts, file="./results/coinfection_normalized_counts_DESeq2.csv")
```
```{r}
rld <- rlog(dds, blind=TRUE)
plotPCA(rld, intgroup="sampletype")
pdf("./results/PlotPCA_dds.pdf")
plotPCA(rld, intgroup="sampletype")
dev.off()
```

```{r}
rld_mat <- assay(rld)
rld_cor <- cor(rld_mat) 
head(rld_cor)
head(meta)
```

```{r}
library(pheatmap)
pheatmap(rld_cor, annotation = meta)

heat.colors <- RColorBrewer::brewer.pal(6, "Blues")
pheatmap(rld_cor, annotation = meta, color = heat.colors, border_color=NA, fontsize = 10, 
         fontsize_row = 10, height=20)

pdf("./results/PlotHeatmap_dds.pdf")
heat.colors <- RColorBrewer::brewer.pal(6, "Blues")
pheatmap(rld_cor, annotation = meta, color = heat.colors, border_color=NA, fontsize = 10, 
         fontsize_row = 10, height=20)
```

```{r}
library(edgeR)
options(digits=3)
infection.targets<-read.delim("./data/fileDesc.txt")
infection.targets
rownames(infection.targets)<-c("Ha1","Ha2","Ha3","Ctr1","Ctr2","Ctr3")
infection.targets
infection <- readDGE(infection.targets, header=F)
dim(infection)
```

```{r}
head(infection)
infection.rawCount <- infection$count
head(infection.rawCount)
```
```{r}
library(ggplot2)

ggplot(infection.rawCount) + geom_histogram(aes(x = Ha1), stat = "bin", bins = 200) + xlab("Raw expression counts") + ylab("Number of genes")

ggplot(infection.rawCount) + geom_histogram(aes(x = Ha1), stat = "bin", bins = 200) + xlab("Raw expression counts") + ylab("Number of genes")

write.csv(infection.rawCount, file="./results/infection.rawCounts.csv")

infection.normCPM <- cpm(calcNormFactors(infection))
dim(infection.normCPM)

head(infection.normCPM)
write.csv(infection.normCPM, file="./results/infection.normCPM.csv")

infection.filtered <- rowSums(cpm(infection)>1) >=3

table(infection.filtered)
```
```{r}
infection$samples$lib.size
Infection <- infection[infection.filtered,]
colSums(Infection$counts)
dim(Infection)
Infection$samples$lib.size <- colSums(Infection$counts)
Infection$samples
Infection = calcNormFactors(Infection)
Infection$samples
Infection.filtered.normCPM <-cpm(calcNormFactors(Infection))
write.csv(Infection.filtered.normCPM, file="./results/Infection.filtered.normCPM.csv")
```

```{r}
group<-factor(c('Ha','Ha','Ha',"Ctr","Ctr","Ctr"))

Infection.design <- model.matrix(~group)   
rownames(Infection.design)<-colnames(Infection$counts)
Infection.design

plotMDS(Infection, main="MDS plot of RNA-Seq", labels=colnames(Infection$counts))

plotMDS(Infection, main="MDS plot of Infection RNA-Seq", labels=colnames(Infection$counts))
```

```{r}
Infection <- estimateGLMCommonDisp(Infection, Infection.design)
Infection <- estimateGLMTrendedDisp(Infection, Infection.design)
Infection <- estimateGLMTagwiseDisp(Infection, Infection.design)
plotMeanVar(Infection, show.tagwise.vars=T,NBline=T)

plotBCV(Infection)

Infection.fit <- glmFit(Infection, Infection.design)
colnames(Infection.fit)

lrt.Ha_vs_Ctr <- glmLRT(Infection.fit, coef=2)  # to compare Ha vs Ctr (Ha_vs_Ctr)

t1<-topTags(lrt.Ha_vs_Ctr, n=nrow(Infection))
```

```{r}
head(t1$table)

summary(decideTests(lrt.Ha_vs_Ctr, adjust.method="BH", p.value=0.05))

nrow(subset(topTags(lrt.Ha_vs_Ctr, n=586)$table,  logFC > 0))

lrt.Ha_vs_Ctr_UP <- subset(topTags(lrt.Ha_vs_Ctr, n=586)$table, logFC > 0)

nrow(subset(topTags(lrt.Ha_vs_Ctr, n=586)$table,  logFC < 0))
lrt.Ha_vs_Ctr_DW <- subset(topTags(lrt.Ha_vs_Ctr, n=586)$table, logFC < 0)
DEtags.lrt.Ha_vs_Ctr <- rownames(Infection)[as.logical(decideTests(lrt.Ha_vs_Ctr, adjust.method="BH", p.value=0.05))]

write.csv(lrt.Ha_vs_Ctr_UP, file="./results/lrt.Ha_vs_Ctr_UP.csv")
write.csv(lrt.Ha_vs_Ctr_DW, file="./results/lrt.Ha_vs_Ctr_DW.csv")
```

```{r}

Infection.colHavsCtr = rep('grey55', nrow(Infection))

Infection.colHavsCtr[lrt.Ha_vs_Ctr$table$PValue < 0.05 & lrt.Ha_vs_Ctr$table$logFC >0 ] <- "red"
Infection.colHavsCtr[lrt.Ha_vs_Ctr$table$PValue < 0.05 & lrt.Ha_vs_Ctr$table$logFC <0 ] <- "blue"

par(omi=c(0.1,0.1,0.1,0.1), las=1, cex=0.5, mgp=c(3,1,0), cex.main=1.8, cex.lab=1.4, cex.axis=1.4)
plotSmear(lrt.Ha_vs_Ctr, de.tags=DEtags.lrt.Ha_vs_Ctr, xlab="log-counts per million (logCPM)", ylab="log2-fold change (log2FC)", main="Ha infection compared to Control", pch=19, cex=0.4, smearWidth=0.5, panel.first=grid(), smooth.scatter=FALSE, ylim=c(-7,7), yaxs="i")

abline(h=c(-1,1),col="dodgerblue")
```

```{r}
par(omi=c(0.1,0.1,0.1,0.1), las=1, cex=0.5, mgp=c(3,1,0), cex.main=1.8, cex.lab=1.4, cex.axis=1.4)
plotSmear(lrt.Ha_vs_Ctr, xlab="log-counts per million (logCPM)", ylab="log2-fold change (log2FC)", main="a infection compared to Control", smearWidth=0.5, pch=21, cex=0.4, deCol="red", col=Infection.colHavsCtr, ylim=c(-7,7), yaxs="i")

abline(h=c(-1,1),col="dodgerblue")
```

```{r}
par(omi=c(0.1,0.1,0.1,0.1), las=1, cex=0.5, mgp=c(3,1,0), cex.main=1.8, cex.lab=1.4, cex.axis=1.4)
plotSmear(lrt.Ha_vs_Ctr, xlab="log-counts per million (logCPM)", ylab="log2-fold change (log2FC)", main="Ha infection compared to Control", smearWidth=0.5, pch=21, cex=0.4, deCol="red", col=Infection.colHavsCtr, ylim=c(-7,7), yaxs="i")
abline(h=c(-1,1),col="dodgerblue")
```
```

